\chapter{Methods}
So in the last two chapters we have looked at looping, making decisions and sequences and we started work on our games.\paragraph{}
So we left the game
\begin{lstlisting}
room="hall“
while True:
  room=input("Which room? ")
  if room=="hall" or room=="Hall":
     print("you are in the hall - there are two rooms off this\n -A kitchen\n-A study\n")
  elif room=="kitchen":
      print("your are in the kitchen")
  elif room=="study": 
     print("you are in the Study")
  else:
      print("You have entered the Twilight Zone and are lost in space and Time")

\end{lstlisting}
It is getting better but we still have some issues.\paragraph{}
We can jump from kitchen to study without going through the hall.
What if we want to have rooms coming of the kitchen to other rooms?
Just a few issues.
\section{define it}
So what if we made the rooms it a ‘thing’ and prints messages but also helps us structure what happens next. So what if we start defining what each room is. I am going to sneek in inputs as well. So we have a way of defining the room.
\begin{lstlisting}
def hall():
  print("you are in the hall - there are two rooms off this\n -A kitchen\n-A study\n")
  room=input("Which room? ")
  if room=="kitchen":
    kitchen()
  elif room=="study":
    study()
  else:
    print("You have entered the Twilight Zone and are lost in space and Time")

def kitchen():
  print("your are in the kitchen, there is nothing in here") 
  print("You leave this room and back to the Hall")
  hall()  

def study():
  print("you are in the Study")
  print("Back to the hall")
  hall()
  
while True:
    hall()
\end{lstlisting}
You always start in the hall
When you leave the kitchen and study you must go back to the hall - this is has just been done to make the changes easier for the moment.\paragraph{}
So what have we done with these def lines is define what happens when we are in a room. We can change what happens in a room without necessarily altering what happens in other rooms. So if I looking for what happens in the kitchen I only need to go and look at:
\begin{lstlisting}
def kitchen():
  print("your are in the kitchen, there is nothing in here") 
  print("You leave this room and back to the Hall")
  hall() 
\end{lstlisting}
These defs are subroutines.

\section{local and global}
So lets add in an item so we can pick it up or put it down.

\begin{lstlisting}
item ="No items"

def hall():
  print("you are in the hall - there are two rooms off this\n -A kitchen\n-A study\n")
  room=input("Which room? ")
  if room=="kitchen":
    kitchen()
  elif room=="study":
    study()
  else:
    print("You have entered the Twilight Zone and are lost in space and Time")

def kitchen():
  #global variable which means change it here change it everywhere
  global item
  item="a knife"
  #when printing out strings we can use the + to
  #combine them and # is for comments
  print("your are in the kitchen, there is "+item) 
  print("You leave this room and back to the Hall")
  hall()  

def study():
  print("You are carrying "+item)
  print("you are in the Study")
  print("Back to the hall")
  hall()
  
while True:
    hall()
\end{lstlisting}
If we had created item inside one of the defs it could only we seen within the define not outside of it. So we need in this case to have a variable item that is seen and can be changed anywhere that is the global bit seen in kitchen().

\subsection{How a go yourself}
Using the Murder Mystery Games build your own and then:
\begin{itemize}
    \item Add new descriptions for all the rooms
    \item Add at least two new rooms.
\end{itemize}

\section{Digging a bit deeper into def}
We saw a bit earlier that we create using def small routines that we can use over and over again. As examples hall() and kitchen(). These we can think of as subroutines (sometimes methods, functions, procedures,...). That allow us to define something once and then use it over and over again.\paragraph{}
So the question arises why bother? There are many positive reasons to do this:
\begin{itemize}
    \item We only need write the code once and so if we want to change what it does we just go to one place and change it and it is changed where it is used.
    \item Related to the last point if something goes wrong with that function we only need to go to the def part that defines it.
    
\end{itemize}
So the elements in the definition of a function can be seen in the example below:
\begin{lstlisting}
  def move(direction):
    if direction == "North"
        return 1
    else:
        print("You can't go that way")
        return 0
\end{lstlisting}
Might look a bit scary but it isn't if we break it down.\paragraph{}
We give the subroutine a name in this case move and we can also allow a value/"thing" to brought into the subroutine in this case direction. In all case that we use the def we end the line with a colon. All the next lines of code that are part of the subroutine are indented just like we saw with loops and ifs.\paragraph{}
Just as we can pass things into the subroutine we can pass things out - this is the role of return. So in the example if direction = "North" a 1 is return otherwise (or else) we get a message and 0 is passed out of the subroutine.\paragraph{}
In the code when we want to use it we could have a line that says 
\begin{lstlisting}
  value=move("East")
\end{lstlisting}

We get a 1 or 0 stored in value depending on what was included in the move().